dispRity palaeo demo

### 2016-05-29 doc/dispRity-ecology-demo.Rmd
Manuals formating

This demo aims to give quick overview of the <span>`dispRity`
</span> package (v.0.2) for ecological analysis. Please refer to [GitHub
page:
github.com/TGuillerme/dispRity](https://github.com/TGuillerme/dispRity)
for other vignettes, namely the `dispRity` tutorial that
explains the functions in more detail.

@@

You can install this package easily if you are using the latest version
of <span>`R` </span>and `devtools`.

@@

For more details on the `dispRity` objects, see the
[<span>`dispRity`
</span>manual](https://github.com/TGuillerme/dispRity/blob/master/doc/dispRity-manual.pdf).
Basically the idea is to avoid jamming the <span>`R` </span>console such
as when using:

@@

Disparity can be calculated in many ways, and therefore, the
`dispRity` function allows to measure disparity *as defined
by the user*. For more details on disparity, see the [other
vignettes](https://github.com/TGuillerme/dispRity).

>>>>

This demo aims to give quick overview of the `dispRity` package (v.0.2) for ecological analysis.
Please refer to [GitHub page](https://github.com/TGuillerme/dispRity) for other vignettes, namely the `dispRity` manual that explains the functions in more details.

@@

You can install this package easily if you are using the latest version of `R` (> 3.0.0) and `devtools`.

@@

For more details on the `dispRity` objects, see the [`dispRity` manual](https://github.com/TGuillerme/dispRity/blob/master/doc/dispRity-manual.pdf).
Basically the idea is to avoid jamming the `R` console such as when using:

@@

Disparity can be calculated in many ways, and therefore, the `dispRity` function allows to measure disparity *as defined by the user*.
For more details on measuring disparity, see the [`dispRity` metric vignette](https://github.com/TGuillerme/dispRity/blob/master/doc/dispRity-metrics.Rmd).

### 2016-05-29 doc/dispRity-manual.Rmd
Manuals formating

-   **Ordinated space**: it designates the mathematical multidimensional
    object studied here. In morphometrics, this one is often referred as
    being the morphospace. However it can also be the the cladisto-space
    for cladistic data or the eco-space in ecology, etc. In practice,
    this term designates an ordinated matrix where the columns represent
    the dimensions of the ordinated space (often – but not necessarily –
    $>$ 3!) and the rows represent the elements within this space.

@@

You can install this package easily if you use the latest version of
<span>`R` </span>and `devtools`.

@@

In this tutorial we are going to use a subset of the ordinated cladistic
data from @beckancient2014 that contains 50 taxa (elements) ordinated
using their cladistic distance (i.e. the distance between their discrete
morphological characters). Note that this data is more oriented towards
palaebiology analysis but that it can apply to other disciplines. Please
refer to the [GitHub page:
github.com/TGuillerme/dispRity](https://github.com/TGuillerme/dispRity)
for other vignettes covering some specific example.

@@

Disparity analysis can involve a lot of shuffling around with many
matrices (especially when bootstrapping the data) which can be
impractical to visualise and will quickly jam your <span>`R`
</span>console. For example, we can have a look at the structure of the
object created in the quick example:

@@

## Summarising it using the S3 method print.dispRity:
median_centroids
## This displays some basic information about the object content
@

\noindent Note however, that it is always possible to recall the full object using the argument \texttt{all=TRUE}:

<<dispRity class3, eval=FALSE>>=
## Displaying the full object
print(median_centroids, all = TRUE)
## But that's 155091 lines!
```

@@

This package aims to be a modular package where users can personalise
some aspects of the package fairly easily. The modular aspect allows the
package users to use several implemented tools to help them creating
personalised functions. For example, the <span>`dispRity`
</span>function intake a `metric` argument designating how disparity
should be calculated. This argument can take some already implemented
functions such as `mean` but also some completely new ones such as the
sum divided by length:

@@

For more information, please refer to the [GitHub page:
github.com/TGuillerme/dispRity](https://github.com/TGuillerme/dispRity)
for the vignette about the metric implementation in <span>`dispRity`
</span>.

@@

In the following section, different functionalities are detailed. The
functions are presented in a “classic” pipelined study fashion for
convenience (i.e. splitting the data, bootstrapping it, calculating
disparity, testing differences). Note however that most of these
functions can be used in any order and that some steps can of course be
skipped.

@@

In these examples we used only simple ordinated matrix but of course, it
might be more interesting to directly use the <span>`dispRity`
</span>objects we generated in the steps above. Thus we can calculate
the bootstrapped and rarefied disparity in each sub-sample of the
morphospace. In this example we are going to define disparity as being
the median distance between each element and each dimensions centroids.

@@

This function is a S3 function (`summary.dispRity`) allowing to
summarize the content of `dispRity` objects that contain
disparity calculations. This function intakes a <span>`dispRity`
</span>object plus various options namely the `quantile` values for the
confidence intervals levels; the `cent.tend` for the central tendency to
use for summarising the results and two *visual* options which are
whether to recall the `dispRity` options and how much
digits are wanted in the results.

@@

Finally, the `dispRity` package allows to apply some tests
to the data in order to properly test hypothesis. The function
`test.dispRity` works in a similar way as the <span>`dispRity`
</span>function: it intakes a `dispRity` object (containing
disparity measurements), a `test` and a `comparisons` argument. The
`test` argument can be any statistical or non-statistical test to apply
to the disparity object. It can be common test function (e.g.
`stats::t.test`) or some functions implemented in <span>`dispRity`
</span>(see `dispRity.test`) or even user defined functions. The
`comparisons` argument must indicate the way the test should be applied
to the data:

@@

If you have any extra suggestions or comments on what has already been
developed or will be developed, please send me an email
(<guillert@tcd.ie>) or if you are a GitHub user, directly create an
issue on the [GitHub page
(github.com/TGuillerme/dispRity)](https://github.com/TGuillerme/dispRity).

>>>>

-   **Ordinated space**: it designates the mathematical multidimensional object studied here.
    In morphometrics, this one is often referred as being the morphospace.
    However it can also be the the cladisto-space for cladistic data or the eco-space in ecology, etc.
    In practice, this term designates an ordinated matrix where the columns represent the dimensions of the ordinated space (often – but not necessarily - > 3!) and the rows represent the elements within this space.

@@

You can install this package easily if you use the latest version of `R` (> 3.0.0) and `devtools`.

@@

In this tutorial we are going to use a subset of the ordinated cladistic data from @beckancient2014 that contains 50 taxa (elements) ordinated using their cladistic distance (i.e. the distance between their discrete morphological characters).
Note that this data is more oriented towards palaebiology analysis but that it can apply to other disciplines.
Please refer to the [GitHub page](https://github.com/TGuillerme/dispRity) for other vignettes covering some specific example.

@@

Disparity analysis can involve a lot of shuffling around with many matrices (especially when bootstrapping the data) which can be impractical to visualise and will quickly jam your `R` console.
For example, we can have a look at the structure of the object created in the quick example:

@@

## Summarising it using the S3 method print.dispRity:
median_centroids
## This displays some basic information about the object content
```

Note however, that it is always possible to recall the full object using the argument \texttt{all=TRUE}:

```{r, eval=FALSE}
## Displaying the full object
print(median_centroids, all = TRUE)
## But that's 155091 lines!
```

@@

This package aims to be a modular package where users can personalise some aspects of the package fairly easily.
The modular aspect allows the package users to use several implemented tools to help them creating personalised functions.
For example, the `dispRity` function intake a `metric` argument designating how disparity should be calculated.
This argument can take some already implemented functions such as `mean` but also some completely new ones such as the sum divided by length:

@@

For more information on the metrics, please refer to the [`dispRity` metric manual](https://github.com/TGuillerme/dispRity/blob/master/doc/dispRity-metric.Rmd) for the vignette about the metric implementation in `dispRity`.

@@

In the following section, different functionalities are detailed. The functions are presented in a "classic" pipelined study fashion for convenience (i.e. splitting the data, bootstrapping it, calculating disparity, testing differences).
Note however that most of these functions can be used in any order and that some steps can of course be skipped.

@@

In these examples we used only simple ordinated matrix but of course, it might be more interesting to directly use the `dispRity` objects we generated in the steps above.
Thus we can calculate the bootstrapped and rarefied disparity in each sub-sample of the morphospace.
In this example we are going to define disparity as being the median distance between each element and each dimensions centroids.

@@

This function is a S3 function (`summary.dispRity`) allowing to summarize the content of `dispRity` objects that contain disparity calculations.
This function intakes a `dispRity` object plus various options namely the `quantile` values for the confidence intervals levels; the `cent.tend` for the central tendency to use for summarising the results and two *visual* options which are whether to recall the `dispRity` options and how much digits are wanted in the results.

@@

Finally, the `dispRity` package allows to apply some tests to the data in order to properly test hypothesis.
The function `test.dispRity` works in a similar way as the `dispRity` function: it intakes a `dispRity` object (containing disparity measurements), a `test` and a `comparisons` argument.
The `test` argument can be any statistical or non-statistical test to apply to the disparity object. It can be common test function (e.g. `stats::t.test`) or some functions implemented in `dispRity` (see `dispRity.test`) or even user defined functions. The `comparisons` argument must indicate the way the test should be applied to the data:

@@

If you have any extra suggestions or comments on what has already been developed or will be developed, please send me an email (<guillert@tcd.ie>) or if you are a GitHub user, directly create an issue on the [GitHub page](https://github.com/TGuillerme/dispRity).

### 2016-05-29 doc/dispRity-palaeo-demo.Rmd
Manuals formating

This demo aims to give quick overview of the <span>`dispRity`
</span>package (v.0.2) for palaeobiology analysis. Please refer to
[GitHub page:
github.com/TGuillerme/dispRity](https://github.com/TGuillerme/dispRity)
for other vignettes, namely the [<span>`dispRity`
</span>manual](https://github.com/TGuillerme/dispRity/blob/master/doc/dispRity-manual.pdf)
that explains the functions in more detail.

@@

You can install this package easily if you are using the latest version
of <span>`R` </span>and `devtools`.

@@

The output object is a `dispRity` object. For details about
this object class, please refer to the [<span>`dispRity`
</span>manual](https://github.com/TGuillerme/dispRity/blob/master/doc/dispRity-manual.pdf).
In brief, `dispRity` objects are lists of different
elements (i.e. disparity results, morphospace time seriess, etc...) that
display only a summary of the object when calling the object for
avoiding jamming the <span>`R` </span>console.

@@

Once we obtain our different time series, we can bootstrap and rarefy
them (i.e. pseudo-replicating the data). The bootstrapping will allow to
make each series more robust to outliers and the rarefaction will allow
us to compare slices with the same number of taxa to get rid of eventual
sampling problems (i.e. more taxa in one bin than the other). The
`boot.matrix` function that allows to bootstrap the <span>`dispRity`
</span>object and the `rarefaction` option within allows to rarefy it.

@@

We can now calculate the disparity within each time series along side
with some confidence intervals generated by the pseudo-replication step
above (bootstraps/rarefaction). Disparity can be calculated in many ways
and this package allows user to come up with their own disparity
metrics. For more info, please refer to the [<span>`dispRity`
</span>manual and the metric
vignette](https://github.com/TGuillerme/dispRity).

@@

As for any <span>`R` </span>package, you can refer yourself to each
individual function for more details.

>>>>

This demo aims to give quick overview of the `dispRity` package (v.0.2) for palaeobiology analysis. Please refer to [GitHub page](https://github.com/TGuillerme/dispRity) for other vignettes, namely the [`dispRity` manual](https://github.com/TGuillerme/dispRity/blob/master/doc/dispRity-manual.Rmd) that explains the functions in more detail.

@@

You can install this package easily if you are using the latest version of `R` and `devtools`.

@@

The output object is a `dispRity` object.
For details about this object class, please refer to the [`dispRity` manual](https://github.com/TGuillerme/dispRity/blob/master/doc/dispRity-manual.Rmd).
In brief, `dispRity` objects are lists of different elements (i.e. disparity results, morphospace time seriess, etc...) that display only a summary of the object when calling the object for avoiding jamming the `R` console.

@@

Once we obtain our different time series, we can bootstrap and rarefy them (i.e. pseudo-replicating the data). The bootstrapping will allow to make each series more robust to outliers and the rarefaction will allow us to compare slices with the same number of taxa to get rid of eventual sampling problems (i.e. more taxa in one bin than the other).
The `boot.matrix` function that allows to bootstrap the `dispRity` object and the `rarefaction` option within allows to rarefy it.

@@

We can now calculate the disparity within each time series along side with some confidence intervals generated by the pseudo-replication step above (bootstraps/rarefaction).
Disparity can be calculated in many ways and this package allows user to come up with their own disparity metrics.
For more info, please refer to the [`dispRity` metric vignette](https://github.com/TGuillerme/dispRity/blob/master/doc/dispRity-metrics.Rmd).

@@

As for any `R` package, you can refer yourself to each individual function for more details.
